### 2019-06-12 2-Chapters/2-Chapter/chapter2.tex
Typos, et commentaires Christophe

In this thesis, we only study analytically the mean regret $R_T^{\cA}$ (in Chapters \ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances used in our work.

>>>>

In this thesis, we only study analytically the mean regret $R_T^{\cA}$ (in Chapters~\ref{chapter:5} and \ref{chapter:6}), but we quickly mention other measures of performances used in our work.

### 2019-06-12 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Typos, et commentaires Christophe

\paragraph{Slotted protocol.}
%
For simplicity, and for consistency with the rest of this thesis, we only consider a discrete protocol (\ie, slotted).
As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
All devices share a synchronized time, and know in advance $K$, the finite number of available RF channels.
In each time slot, the devices try to send packets to the unique gateway, which listens continuously to all channels, following an ALOHA-based communication, with no sensing.
Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the gateway.
In one channel, if only one packet is sent in this part of the slot, the gateway can decode it and sends an acknowledgement to the device in the second part (on the same channel).
If two or more devices send an uplink packet in the same slot, the uplink packets collide, and the acknowledgement (\Ack) is not transmitted. In this case, we say that there is a \emph{collision} in this time slot.
In other words, if the gateway cannot successfully decode the incoming (uplink) messages because they were corrupted by collisions, it does not send any \Ack{} back.
This way, no collision can occur on the downlink messages, easing the analysis of collisions.

@@

\paragraph{First results.}
%
Figure \ref{fig:41:from10to100} presents the successful transmission rate as a function of time.
The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy (which are outperformed easily by the MAB algorithms), and the two (optimal and greedy) oracle policies (which outperform slightly the MAB algorithms).
The results are displayed when $10\%$, $30\%$, $50\%$ and $100\%$ of the traffic is generated by dynamic devices.

We can see in Figure \ref{fig:41:from10to100} that the TS algorithm (\textcolor{red}{in red}) outperforms the \UCB{} algorithm (\textcolor{blue}{in blue}), when the number of end-devices is below 50\%. When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after a small number of transmissions (\ie, they show quick convergence).
Moreover, we can see in Figures \ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the optimal oracle and suboptimal oracle policies.
For instance, for $10\%$ of dynamic devices, after about $1000$ transmissions, using \UCB{} over the naive uniform policy improved the successful transmission rate from $83\%$ to $88\%$, and using Thompson Sampling improved it to $89\%$.
Increasing the number of end-devices decreases the gap between the optimal and random policies:
as expected intuitively, the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures~\ref{fig:41:100intelligent} and on the right end side of Figure~\ref{fig:41:perf_learning}.

@@

Figure \ref{fig:41:perf_learning} also shows that learning keeps near-optimal performance even when the proportion of devices becomes large.
Note that when this proportion increases, the assumptions of a stochastic MAB model are clearly violated, and there is no mathematical justification for the efficiency of TS and \UCB{} algorithms.
Hence, it is surprising to have near optimal performance with stochastic MAB algorithms applied to partly or fully dynamic scenarios.

>>>>

\paragraph{Slotted protocol.}
%
For simplicity, and for consistency with the rest of this thesis, we only consider a discrete protocol (\ie, slotted).
As illustrated in Figure~\ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
All devices share a synchronized time, and know in advance $K$, the finite number of available RF channels.
In each time slot, the devices try to send packets to the unique gateway, which listens continuously to all channels, following an ALOHA-based communication, with no sensing.
Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the gateway.
In one channel, if only one packet is sent in this part of the slot, the gateway can decode it and sends an acknowledgement to the device in the second part (on the same channel).
If two or more devices send an uplink packet in the same slot, the uplink packets collide, and the acknowledgement (\Ack) is not transmitted. In this case, we say that there is a \emph{collision} in this time slot.
In other words, if the gateway cannot successfully decode the incoming (uplink) messages because they were corrupted by collisions, it does not send any \Ack{} back.
This way, no collision can occur on the downlink messages, easing the analysis of collisions.

@@

\paragraph{First results.}
%
Figure~\ref{fig:41:from10to100} presents the successful transmission rate as a function of time.
The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy (which are outperformed easily by the MAB algorithms), and the two (optimal and greedy) oracle policies (which outperform slightly the MAB algorithms).
The results are displayed when $10\%$, $30\%$, $50\%$ and $100\%$ of the traffic is generated by dynamic devices.

We can see in Figure~\ref{fig:41:from10to100} that the TS algorithm (\textcolor{red}{in red}) outperforms the \UCB{} algorithm (\textcolor{blue}{in blue}), when the number of end-devices is below 50\%. When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after a small number of transmissions (\ie, they show quick convergence).
Moreover, we can see in Figures~\ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the optimal oracle and suboptimal oracle policies.
For instance, for $10\%$ of dynamic devices, after about $1000$ transmissions, using \UCB{} over the naive uniform policy improved the successful transmission rate from $83\%$ to $88\%$, and using Thompson Sampling improved it to $89\%$.
Increasing the number of end-devices decreases the gap between the optimal and random policies:
as expected intuitively, the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures~\ref{fig:41:100intelligent} and on the right end side of Figure~\ref{fig:41:perf_learning}.

@@

Figure~\ref{fig:41:perf_learning} also shows that learning keeps near-optimal performance even when the proportion of devices becomes large.
Note that when this proportion increases, the assumptions of a stochastic MAB model are clearly violated, and there is no mathematical justification for the efficiency of TS and \UCB{} algorithms.
Hence, it is surprising to have near optimal performance with stochastic MAB algorithms applied to partly or fully dynamic scenarios.

### 2019-06-12 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Typos, et commentaires Christophe

We simulate our network considering $N$ devices following the contention Markov process described in Section \ref{sub:43:model}, and a LoRaWAN standard with $K=4$ channels, as in Section~\ref{sec:4:firstModel}.
Each device is set to transmit with a fixed probability $p=10^{-3}$, \ie, a packet about every $20$ minutes for time slots of $1\;\mathrm{s}$.
%
For the evaluation of the proposed heuristics, a total number of $T=20 \times 10^4$ time slots is considered, and the results are averaged over $1000$ independent random simulations.
%In this way, it allows us to evaluate the learning rate, as well as the convergence of each learning approach.

>>>>

We simulate our network considering $N$ devices following the contention Markov process described in Section~\ref{sub:43:model}, and a LoRaWAN standard with $K=4$ channels, as in Section~\ref{sec:4:firstModel}.
Each device is set to transmit with a fixed probability $p=10^{-3}$, \ie, a packet about every $20$ minutes for time slots of $1\;\mathrm{s}$.
%
For the evaluation of the proposed heuristics, a total number of $T=20 \times 10^4$ time slots is considered, and the results are averaged over $1000$ independent random simulations.
%In this way, it allows us to evaluate the learning rate, as well as the convergence of each learning approach.
